Domain-Driven, Generative Data Model for BigPetStore

### 2014-07-23 bdcloud_paper/latex/paper.tex
rewrote intro and abstract

\begin{abstract}
Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key for consistent development, testing, and benchmarking of big data software, algorithms, and infrastructure. BigPetStore is a big data application blueprint, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics, built around a fictional chain of pet store which has been used for testing and benchmarking in the Apache BigTop distribution. We present a new generative data model for BigPetStore and demonstrate the model's ability to generate semantically-rich data that can be used for complex analytics workflows at variable scale ranging from a single machine to a large cluster.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

\section{Introduction}
Big data applications process large, continuously-changing data sets with the general goal of information and knowledge extraction.  New tools such as Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Shark \cite{Engle2012} based around big data frameworks such as Hadoop \cite{Hadoop}, Storm \cite{Storm}, and Spark \cite{Zaharia2010,Zaharia2012} are being released every day. Development, testing and benchmarking of big data software, algorithms, and infrastructure pose unique challenges.  Though software and algorithms will be deployed to large clusters for benchmarking and production usage, engineers develop and test these tools on desktops due to the speed and ease of development and debugging on a local machine and limited access to large clusters due to the high costs of purchasing, provisioning, and maintaining such clusters.  Software and algorithms are being developed at incredible speed with new releases everyday. The high speed of development driven by the need to be the first to market in a competitive field often results in buggy software.  There is great need for consistent and rigorous testing and benchmarking frameworks to aid in the identification of bugs while maintaining the high-speed of development.  Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key to increasing access to and development of consistent and rigorous testing and benchmarking frameworks.

Over time, more high-quality data sets are becoming publicly available, but these data sets often carry burdens that force developers to make undesirable trade offs. Since few data sets are scalable (allowing the user to pick how much data they want), users are forced to choose between small data sets which are useful for development and testing on a local machine, but not for production-scale tests or benchmarking, or large data sets which are prohibitively expensive to host, transfer, and store. Furthermore, access to large data sets can be unreliable due to dependence on third party benefactors for hosting and requires significant amounts of bandwidth to download.  Several companies (such as Twitter\cite{Twitter}) offer access to their data but the costs and usage restrictions can be undesirable.

Methods for generating synthetic data offer multiple advantages over pre-generated data sets.  The code for generating the data is often small, making it easy and cheap to host, transfer, and store. Licensing restrictions and data privacy issues are avoided with synthetic data sets.  And, the size of the data sets can be scaled arbitrarily, enabling seamless usage of the same data set across a range of situations.

Several synthetic data set generators exist.  TeraGen and the Intel Hadoop Benchmark Suite \cite{Huang2010} can generate data sets quickly and at any scale. Targeted towards specific benchmarks such as Map-Reduce jobs and Hive queries, the resulting data is semantically-void and not useful for other tasks.  Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}. Several approaches use domain-specific languages \cite{Bruno2005,Hoag2007} to allow users to specify additional constraints such as which distributions to sample from or allow modeling basic relationships between fields such as simple arithmetic equations \cite{Alexandrov2012}. Arasu, et al. \cite{Arasu2011} demonstrated that constraint-solving techniques could be used as an alternative to procedural approaches. These frameworks allow for reproducing the structural properties of real data, but the frameworks are not expressive enough to describe the dynamic generation processes and latent variables that would be needed to embed the desired informational complexity and rich semantics needed for analytics.

Realizing the difficulty in creating a generic framework capable of modeling the semantics of deep domain knowledge, recent work\cite{Alexandrov2013} has focused on generating synthetic data sets that satisfy characteristics learned from a real reference data sets. Such approaches appear promising but will need to overcome the difficulties of accurately training models, especially on raw data instead of features.

Until such methods have reached maturity, the development of customized, domain-driven models is best suited for satisfying the needs of big data software and algorithms development. Several customized, domain-driven models have seen usage in real-world situations.  Information Discovery and Analysis System (IDAS) Data Set Generator (IDSG) \cite{Jeske2005,Lin2006} was developed for the Department of Homeland Security to develop synthetic data for effectiveness of IDASs designed to identify potential terrorism threat scenarios.

\textcolor{red}{Huppler, Pavlo}.  \textcolor{red}{standard benchmarks}

The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout \cite{Mahout} from across the Hadoop and Spark ecosystems. BigPetStore's usefulness has been demonstrated by its incorporation into the open-source Apache BigTop distribution \cite{BigTop} for testing and as a reference application.

In this work, we present a new domain-driven model and simulation for generating arbitrarily large, complex data for testing and benchmarking of algorithms, software, and infrastructure in the big data space. Compared with BigPetStore's previous data generator, our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter.  Combining the features of the TeraGen (scalability)  and MovieLens \cite{MovieLens} (semantically rich) input data sets commonly used for big data benchmarking, we present a scalable synthetic data set generation implementation which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.  To validate this approach, we perform an example analysis designed to guide a fictional advertising campaign (a typical business use case) to demonstrate that the model generates data with desirable properties (global patterns and region specific transactions).  We have made the model's implementation available as open source software.

@@

The proposed transaction time is found from the exhaustion times (Eq.~\ref{eq:transaction-times}).  For each item category, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an exponential distribution parameterized separately for each customer $c$. A proposed transaction time $T_{P, i}$ for each item category is found by subtracting the offset time $T_{O, i}$ from the exhaustion time $T_{E, i}$.  The earliest proposed transaction time is taken as the overall proposed transaction time $T_T$.

@@

The item category for each purchase is chosen by sampling from the item categories with the probability of each item category $i$ proportionate to its weight $w_i$.

>>>>

\begin{abstract}
Both developers and users of big data systems benefit from realistic example applications.  For developers, such as applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.

We present a new domain-driven, generative data model for BigPetStore, a big data application blueprint for the Hadoop ecosystem and included in the Apache BigTop distribution. We describe the model and demonstrate the its ability to generate semantically-rich data at variable scale ranging from a single machine to a large cluster.  We validate the model by using the generated data to answer questions about customer locations and purchasing habits for a fictional targeted advertising campaign, a common business use case.

%Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key for consistent development, testing, and benchmarking of big data software, algorithms, and infrastructure. BigPetStore is a big data application blueprint, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics, built around a fictional chain of pet store which has been used for testing and benchmarking in the Apache BigTop distribution. We present a new generative data model for BigPetStore and demonstrate the model's ability to generate semantically-rich data that can be used for complex analytics workflows at variable scale ranging from a single machine to a large cluster.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

\section{Introduction}
Big data applications process large, continuously-changing data sets with the general goal of information and knowledge extraction.  With the wide variety of big data tools available and lagging documentation, both developers and users of big data systems benefit from realistic example applications.  For developers, such as applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.

BigPetStore is a big data application blueprint built around processing transaction data for a fictional chain of pet stores.  BigPetStore mainly targets the Hadoop \cite{Hadoop} ecosystem with examples implemented for loading, cleaning, aggregating and performing analytics on data using Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Mahout \cite{Mahout}. BigPetStore has been incorporated with the open-source Apache BigTop distribution \cite{BigTop}, where it is used for testing and as a reference application. One of BigPetStore's key features is the ability to scale from a single machine to a large cluster, making it easy to develop projects on a local machine and transfer the application to a large cluster for production testing.

To achieve the project's goals of providing high-quality, realistic examples, BigPetStore requires semantically-rich, complex data. At its core, BigPetStore relies on a generative data model for producing synthetic transaction data. Despite the growing number of real data sets now publically available, synthetic data generators have a number of advantages for BigPetStore over real data sets. The synthetic data generator can be packaged with BigPetStore, avoiding the cost, time, and infrastructure needed to host, transfer, and store large data sets. Synthetic data generators are scalable, allowing the user to choose how much data to generate -- a requirement for supporting BigPetStore's goal of running on both single machines and large clusters.  Licensing and privacy issues are avoided with the use of synthetic data sets. And lastly, synthetic data generators are tunable by the user, allowing the user to generate data with specific criteria amenable to testing.  For example, a user may generate transactions from a small number of purchasable items so that clustering results can be easily visualized.

% New tools such as Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Shark \cite{Engle2012} based around big data frameworks such as Hadoop \cite{Hadoop}, Storm \cite{Storm}, and Spark \cite{Zaharia2010,Zaharia2012} are being released every day. Development, testing and benchmarking of big data software, algorithms, and infrastructure pose unique challenges.  Though software and algorithms will be deployed to large clusters for benchmarking and production usage, engineers develop and test these tools on desktops due to the speed and ease of development and debugging on a local machine and limited access to large clusters due to the high costs of purchasing, provisioning, and maintaining such clusters.  Software and algorithms are being developed at incredible speed with new releases everyday. The high speed of development driven by the need to be the first to market in a competitive field often results in buggy software.  There is great need for consistent and rigorous testing and benchmarking frameworks to aid in the identification of bugs while maintaining the high-speed of development.  Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key to increasing access to and development of consistent and rigorous testing and benchmarking frameworks.

%Over time, more high-quality data sets are becoming publicly available, but these data sets often carry burdens that force developers to make undesirable trade offs. Since few data sets are scalable (allowing the user to pick how much data they want), users are forced to choose between small data sets which are useful for development and testing on a local machine, but not for production-scale tests or benchmarking, or large data sets which are prohibitively expensive to host, transfer, and store. Furthermore, access to large data sets can be unreliable due to dependence on third party benefactors for hosting and requires significant amounts of bandwidth to download.  Several companies (such as Twitter\cite{Twitter}) offer access to their data but the costs and usage restrictions can be undesirable.

%Methods for generating synthetic data offer multiple advantages over pre-generated data sets.  The code for generating the data is often small, making it easy and cheap to host, transfer, and store. Licensing restrictions and data privacy issues are avoided with synthetic data sets.  And, the size of the data sets can be scaled arbitrarily, enabling seamless usage of the same data set across a range of situations.

A variety of approaches for generating synthetic data set exist.  TeraGen and the Intel Hadoop Benchmark Suite \cite{Huang2010} are popular tools that can generate data sets quickly and at any scale, but the resulting data is semantically-void and not useful for much more than simple benchmarks.  Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}. Several approaches even provide domain-specific languages \cite{Bruno2005,Hoag2007} for specifying additional constraints such as which distributions to sample from and allow for modeling basic relationships between fields using simple arithmetic equations \cite{Alexandrov2012}. Arasu, et al. \cite{Arasu2011} demonstrated that constraint-solving techniques could be used as an alternative to procedural approaches. These frameworks allow for reproducing the structural properties of real data, but the frameworks are not expressive enough to describe the dynamic generation processes and latent variables that would be needed to embed the desired informational complexity and rich semantics needed for BigPetStore's analytics examples.

Realizing the difficulty in creating a generic framework capable of modeling the semantics of real data, recent work\cite{Alexandrov2013} has focused on generating synthetic data sets that satisfy characteristics learned from real reference data. Such approaches appear promising but will need to overcome the difficulties of accurately training models, especially on raw data instead of features.

Until such methods have reached maturity, BigPetStore's needs are best met by a customized, domain-driven model. Another example of such a model is The Information Discovery and Analysis System (IDAS) Data Set Generator (IDSG) \cite{Jeske2005,Lin2006}. IDSG was developed to to test the effectiveness of IDASs in identifying potential terrorism threat scenarios using synthetic data.  Like BigPetStore, the developers of IDSG needed to avoid licensing restrictions and privacy issues associated with real data.

%\textcolor{red}{Huppler, Pavlo}.  \textcolor{red}{standard benchmarks}

%The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout \cite{Mahout} from across the Hadoop and Spark ecosystems. BigPetStore's usefulness has been demonstrated by its incorporation into the open-source Apache BigTop distribution \cite{BigTop} for testing and as a reference application.

BigPetStore's current model has been used successfully to generate terabytes of synthetic data and is used regularly for testing in Apache BigTop, showcasing the value of the approach.  The current model is limited, however, in its ability to generate data that is semantically rich, limiting progress on BigPetStore's goals of providing realistic analytics examples.

In this work, we present a new domain-driven model and simulation method for BigPetStore.  Compared with BigPetStore's previous data generator, our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter.  Combining the features of the TeraGen (scalability)  and MovieLens \cite{MovieLens} (semantically rich) input data sets commonly used for big data benchmarking, we present a scalable synthetic data set generation implementation which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.  To demonstrate that the model generates data with desirable properties, we performed an example analysis designed to guide a fictional advertising campaign (a typical business use case). The model's implementation available as open source software.

@@

The proposed transaction time is found from the exhaustion times (Eq.~\ref{eq:transaction-times}).  For each item category, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an exponential distribution parameterized separately for each customer $c$. A proposed transaction time $T_{P, i}$ for each item category is found by subtracting the offset time $T_{O, i}$ from the exhaustion time $T_{E, i}$.  The earliest proposed transaction time is taken as the overall proposed transaction time $T_T$. \textcolor{red}{$\beta_c$}

@@

The item category for each purchase is chosen by sampling from the item categories with the probability of each item category $i$ proportionate to its weight $w_i$. \textcolor{red}{$\lambda_i$ - maybe these should be per customer / item category instead of just per item category}
